FraCaS: 80 percent complete

### 2019-01-18 paper2018/paper.tex
packages
paper structure
sec. 2

\documentclass{article}
\usepackage{todonotes}

@@

@@

% - Two-phase system (Dyn semantics from theorem proving)
% - Anaphora (as in Jolli paper) --- but this is the first time that this is done for FraCas.
% - Multiple readings
% - Threshold-based interpretation of (some) adjectives + Linear arithmetic proofs
% - New handling of prepositions and adverbs
% - Definites handled by adding an assumption at the top-level. (instead as inline existential) -- Enabled by dyn. semantics
% - Plurals/Quantifiers improved
% - Genders are properties of nouns
% - Handling of comparatives (with numbers)

@@

>>>>

\documentclass{article}
\usepackage{todonotes}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

\usepackage[backend=bibtex,citestyle=authoryear,style=alphabetic,natbib]{biblatex}
\addbibresource{PaperTools/bibtex/jp.bib}
\addbibresource{phorae.bib}

@@

% - Two-phase system (Dyn semantics from theorem proving)
% - Anaphora (as in Jolli paper) --- but this is the first time that this is done for FraCas.
% - Multiple readings
% - Threshold-based interpretation of (some) adjectives + Linear arithmetic proofs
% - New handling of prepositions and adverbs
% - Definites handled by adding an assumption at the top-level. (instead as inline existential) -- Enabled by dyn. semantics
% - Plurals/Quantifiers improved
% - Genders are properties of nouns
% - Handling of comparatives (with threshold BUT not those of adjectives), relying largely on the dyn. semantic system

@@

\section{Introduction}

\paragraph{Contributions}
(see list above)
\section{Background}

\section{Overview of the system}

Our system consists of two main parts.
\begin{enumerate}
\item A converter from syntax trees to types. The syntax trees follow
  the GF formalism, and the types follow the Coq formalism. The
  converter itself is a Haskell program. It implements a dynamic
  semantics. It comprises the bigger part of our system.
\item A number of type-theoretical combinators. These take care of the
  natural semantics which have no influence on the dynamical
  aspects. This includes for example the treatment of adjectives
  (intersective, subsective, etc.) and adverbs (veridical or not).
\end{enumerate}
This architecture is represented schematically in Figure \ref{fig:overview}.

@@

All underlying systems (GF, Haskell, Coq) are based on lambda calculi
with types. We ensure that each translation preserve typing, locally:
\begin{enumerate}
\item Every GF syntactic category $C$ is mapped to a type noted $⟦C⟧$.
\item GF Functional types are mapped compositionally : $⟦A → B⟧ = ⟦A⟧ → ⟦B⟧$
\item Every GF syntactic construction function $f : X$ is mapped to a function $⟦f⟧$ such that $⟦f⟧ : ⟦X⟧$.
\item GF function applications are mapped compositionally: $⟦t(u)⟧ = ⟦t⟧ (⟦u⟧)$.
\end{enumerate}
Because all system embed the simply-typed lambda calculus, ensuring
type-preservation locally means that types are preserved globally.
Therefor, we are certain that every GF syntax tree can be mapped to
Haskell, and eventuall Coq, without error.

The dynamic semantics follows a monadic structure, as pioneered by
\citet{Shan:2002}. There are two kinds of effects carried by the
monad.  The first one comprises a series of state updates and queries,
corresponding to all elements that can be referred to by anaphoric
expressions. These can be usual ones (like NPs), but also less usual
ones (like 2-place verbs, or a quantity --- which we illustrate
below). The other kind of effects is \emph{non-determinism}. This
models the property that linguistic expressions can have several
interpretations. The monadic structure allows to locally express that
a given expression has several meaning; the monadic bind ensures that
all combinations of meanings are considered at the top-level,
combinatorially. This dynamic semantics allows us to model many
phenomena in a precise way:

\paragraph{anaphora} We can handle anaphora % wait so see what JoLLi says here

  % - Dynamic Semantics/Anaphora (as in Jolli paper) --- but this is the first time that this is done for FraCas.
%   -> Gender is a property of nouns
\paragraph{ellipsis} Ellipsis is handled in essentially the same way
as anaphora. This method is made especially straightforward thanks to
using GF syntax tree, which requires an explicit argument for each
predicate. Thus, ellpsis are made explicit by the parsing phase.

\paragraph{definites} A naive way to handle definites is as sigma (or
exsistential types). \todo{add example} However, if the semantics
does not feature a dynamic element, then the existential
quantification is introduced locally. This means that the quantifier
can be introduced in the wrong context. Consider the phrase
``everyone pets the dog''. The structure of the interpretation would
be be $∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.

Instead, our take is that definites should be treated as anaphoric
expressions. If the referent is not found in the discourse, then it
will be introduced using a sigma type, \emph{at the top-level} of
the expression. Therefore we record all definites without referent,
using a monadic effect. When the complete sentence has been
interpreted, we re-introduced referent using existentials. For the
above example, we obtain the desired interpretation:
be be $∃y. dog(y) ∧ ∀x. person(x) → pet(x,y)$.

\paragraph{Phrasal comparatives}
Previous attempts to tackle section of the FraCaS test suite showed
that handling comparatives is not easy. Our strategy here is to
leverage our dynamic semantics, revealing an anaphoric element of
comparatives. Consider the hypothesis of FraCaS 239: ``ITEL won more
orders than APCOM lost.''  We postulate that this sentence is
equivalent to the following two separate parts: ``APCOM lost zero or
more orders. ITEL won more orders [than some elliptic quantity].''
The elliptic quantity is introduced every time we talk about some
quantity (indexed by a CN, in this case ``orders''), and it can be
referred to by a comparative, later in the discourse. Using this idea,
we can go one level deeper in the interpretation of our example:
``APCOM lost θ orders. θ >= 0.  ITEL won at least θ+1 orders.''. We
see here how the quantities are introduced. They are added to the
environment so that, they can be referred to as elliptic quantity
expressions. Finally, ``more'' is intepreted as ``at least <elliptic
quantity>+1''.

\paragraph{Adjectives}
In our system, we interpret gradable adjectives using a pair of a
measure $m : objects → Z$ and and a threshold $τ : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $τ$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 > 0$).
\todo{reference to this view in the litterature}.  Additionally,
adjectival assumptions as given in FraCaS are interpreted as linear
equations in $Z$. Because solving such equations is decidable, solving
such problems can be made efficient. Indeed, the tactic that Coq
offers for this purpose can solve all such problems in FraCaS,
automatically.

\paragraph{Adverbs}
Another point, of minor theoretical importance but major practical
one, is our handling of adverbial phrases. We interpret adverbs (and
in general all adverbial and prepositional phrases) are VP-modifiers:
$Adv = VP -> VP$, where $VP = object -> Prop$. However, applying
adverbs to verb-phrases heavily complicates the Coq proofs, because
such phrases can contain quantifiers. Therefore, we instead move the
adverbs, so that they apply to (atomic) verbs only, and can simplify
proofs accordingly. \todo{theoretical justification in the literature}

\section{Results and evaluation}
% - show overall numbers
% - compare with others
% - error analysis (classify, show examples)

\section{Future work}
\section{Conclusion}

\printbibliography
